---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo=TRUE}
# set working directory 
#setwd('')

# load the dataset
df <- read.csv(unz("activity.zip", "activity.csv"), header = TRUE)

# get the first rows of the dataset
head(df)

# get the last rows of the dataset
tail(df)

## check the structure
str(df)

## get the date as date
df$date <- as.Date(df$date)

```


## What is mean total number of steps taken per day?

```{r, echo=TRUE}
daily_steps <- tapply(df$steps, df$date, sum, na.rm=TRUE, simplify=T)
# remove NA's
daily_steps <- daily_steps[!is.na(daily_steps)]

hist(x=daily_steps,
     col="green",
     breaks=20,
     xlab="Daily Steps Taken (without NA's)",
     ylab="Frequency")
```

Means and median of steps taken
```{r, echo=TRUE}
mean(daily_steps)
median(daily_steps)
```


## What is the average daily activity pattern?
```{r, echo=TRUE}
avg_steps <- tapply(df$steps, df$interval, mean, na.rm=TRUE, simplify=T)
df_avg <- data.frame(interval=as.integer(names(avg_steps)), avg=avg_steps)

with(df_avg,
     plot(interval,
          avg,
          type="l",
          xlab="5-minute intervals",
          ylab="average steps in the interval across all days"))
```


## Imputing missing values
```{r, echo=TRUE}
# seems like there are NA's in the data check out how many (we could do the following with a simple sapply as well)
sum(is.na(df$steps))
sum(is.na(df$date))
sum(is.na(df$interval))

# seems there NA's only in steps. Check the percentage 
sum(is.na(df$steps))/nrow(df) # (gives 13%. We will just leave this here)

```
Let's choose a strategy to impute the missing valus
The strategy we will use is the mean data for each 5-min interval to impute to the NA's in that interval

```{r, echo=TRUE}
Imputed_Df <- df
NAs <- is.na(Imputed_Df$steps)
avg_steps <- tapply(df$steps, df$interval, mean, na.rm=TRUE, simplify=T)
Imputed_Df$steps[NAs] <- avg_steps[as.character(Imputed_Df$interval[NAs])]
```
Now let's make a histogram
```{r, echo=TRUE}
daily_steps_Imputed<- tapply(Imputed_Df$steps, Imputed_Df$date, sum, na.rm=TRUE, simplify=T)
hist(x=daily_steps_Imputed,
     col="green",
     breaks=20,
     xlab="Daily Steps Taken (after imputation)",
     ylab="Frequency")
```

## Are there differences in activity patterns between weekdays and weekends?
Getting the weekdays first 
```{r, echo=TRUE}
library(dplyr)
Imputed_Df$weekdays <- weekdays(Imputed_Df$date)
Imputed_Df <- mutate(Imputed_Df, weekdays_factor = ifelse(weekdays == 'Sonntag' | weekdays == 'Samstag', 'weekend', 'weekday' ))
```

Now create a plot

```{r, echo=TRUE}
df_WeekdaysAVG <- aggregate(steps ~ weekdays_factor + interval, data=Imputed_Df, na.rm = TRUE, FUN=mean)

library(lattice)
xyplot(steps ~ interval | factor(weekdays_factor),
       layout = c(1, 2),
       xlab="Interval",
       ylab="Number of steps taken",
       type="l",
       lty=1,
       data=df_WeekdaysAVG)
```